### 2018-03-13 paper/3-compute.tex
Bring fetch back

Our first key idea is the task-description abstraction:
\begin{minted}[xleftmargin=10pt]{haskell}
type Task c k v = @\std{forall}@ f. c f => (k -> f v) -> k -> Maybe (f v)
\end{minted}
This highly-abstracted type is best introduced through an example.
Consider this Excel spreadsheet (c.f. \S\ref{sec-background-excel}):
\vspace{1mm}
\begin{minted}[xleftmargin=10pt]{bash}
  A1: 10     B1: =A1 + A2
  A2: 20     B2: =B1 * 2
\end{minted}
\vspace{1mm}
Here cell \hs{A1} contains the value \hs{10} while cell \hs{B1} contains
the formula \hs{=A1+A2}.
We can represent the formulae of of this spreadsheet
with the following task description
\vspace{1mm}
\begin{minted}[xleftmargin=10pt]{haskell}
sprsh1 :: Compute Applicative String Integer
sprsh1 get "B1" = Just ((+) <$> get "A1" <*> "A2")
sprsh1 get "B2" = Just ((* 2) <$> get "B1")
sprsh1 _   _    = Nothing
\end{minted}
\vspace{1mm}
Here we instantiate keys \hs{k} with \hs{String}, and values \hs{v} with \hs{Integer}.
(Real spreadsheet cells would contain a wider range of values, of course.)
The task description \hs{sprsh1} embodies all the \emph{formulae} of the spreadsheet,
but not the constant values.  Like every \hs{Task}, \hs{sprsh1} is given
a \hs{get} function and a key. It pattern-matches on the key to see if it has
a task description (in the Excel case, a formula)
for it.  If not, it returns \hs{Nothing}, indicating that there is no formula in
that cell.  If there is a formula in the cell, it computes
the value of the formula, using \hs{get} to find the value of any keys on
which it depends.

@@

For now, we content ourselves with observing that a task description,
of type \hs{Task c k v}, is completely isolated from the world of
compilers, calculation chains, file systems, caches, and all other
complexities of real build systems.  It just computes a single output, in
a side-effect-free way, using a call-back (\hs{get}) to get the values
of its dependencies.

>>>>

Our first key idea is the task-description abstraction:
\begin{minted}[xleftmargin=10pt]{haskell}
type Task c k v = @\std{forall}@ f. c f => (k -> f v) -> k -> Maybe (f v)
\end{minted}
This highly-abstracted type is best introduced through an example.
Consider this Excel spreadsheet (c.f. \S\ref{sec-background-excel}):
\vspace{1mm}
\begin{minted}[xleftmargin=10pt]{bash}
  A1: 10     B1: =A1 + A2
  A2: 20     B2: =B1 * 2
\end{minted}
\vspace{1mm}
Here cell \hs{A1} contains the value \hs{10} while cell \hs{B1} contains
the formula \hs{=A1+A2}.
We can represent the formulae of of this spreadsheet
with the following task description
\vspace{1mm}
\begin{minted}[xleftmargin=10pt]{haskell}
sprsh1 :: Compute Applicative String Integer
sprsh1 fetch "B1" = Just ((+)   <$> fetch "A1" <*> "A2")
sprsh1 fetch "B2" = Just ((* 2) <$> fetch "B1")
sprsh1 _   _      = Nothing
\end{minted}
\vspace{1mm}
Here we instantiate keys \hs{k} with \hs{String}, and values \hs{v} with \hs{Integer}.
(Real spreadsheet cells would contain a wider range of values, of course.)
The task description \hs{sprsh1} embodies all the \emph{formulae} of the spreadsheet,
but not the constant values.  Like every \hs{Task}, \hs{sprsh1} is given
a \hs{fetch} function and a key. It pattern-matches on the key to see if it has
a task description (in the Excel case, a formula)
for it.  If not, it returns \hs{Nothing}, indicating that there is no formula in
that cell.  If there is a formula in the cell, it computes
the value of the formula, using \hs{fetch} to find the value of any keys on
which it depends.

@@

For now, we content ourselves with observing that a task description,
of type \hs{Task c k v}, is completely isolated from the world of
compilers, calculation chains, file systems, caches, and all other
complexities of real build systems.  It just computes a single output, in
a side-effect-free way, using a call-back (\hs{fetch}) to obtain the values
of its dependencies.
